*Prochlorococcus marinus* responses to light and oxygen

### 2024-03-01 Code/Manuscript_O2.Rmd
Added Mireille edits

Under 250 µM O~2~ growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white, red or blue LED light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^ growth trends were similar to 250 µM O~2~. SS120 showed no
growth under a 4 h photoperiod for red spectral waveband, however under blue light, SS120 was able to grow under a 4 h photoperiod (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew under 4 and 8 h blue light and 16 h red light photoperiods, however the growth rate decreased under 12 and 16 h white LED light photoperiod treatments. Blue light treatments under PAR of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h photoperiod. The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

@@

The GAM model in Figure \@ref(fig:SS120GAM) summarizes growth responses of SS120 to red (A,B) or blue (C,D) peak PAR and photoperiod, across the 2 imposed oxygen concentrations. Under 250 µM O~2~, Figure \@ref(fig:SS120GAM)C showed highest growth rates below PAR of 50 µmol photons m^-2^ s^-1^ and photoperiods between 8 and 12 h, indicated by the contour line labeled 0.19 d^-1^ (representing the 90^th^ percentile of achieved growth rate).  Under 250 µM O~2~ SS120 is constrained to deeper ocean waters through its intolerance of higher blue PAR levels. These findings align with @mooreComparativePhysiologySynechococcus1995 and are expected for a low light clade.  The disjunct regions of the GAM plot results from variable growth success of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somewhat faster. In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol photons m^-2^ s^-1^, indicated by the contour line labeled 0.4 d^-1^ (representing the 90^th^ percentile of achieved growth rate), out pacing the 90^th^ percentile fastest growth rates under 250 µM O~2~. Under red light and 25 µM O~2~ (Figure \@ref(fig:SS120GAM)B) SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growth under long photoperiods and peak PAR below 50 \~100 µmol photons m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2~]. SS120 did not, however, grow reliably under tested conditions at 2.5 µM O~2~.

@@

Under 250 µM O~2~ growth rates of *P. marinus* MIT9313, Clade LLIV, increased with longer
photoperiods, under low 30 µmol photons m^-2^ s^-1^ peak PAR, (Figure \@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiod. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light at 180 µmol photons m^-2^ s^-1^ peak PAR, under the 8 h photoperiod, consistent with [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

For MIT9313 under 25 µM O~2~ growth rate increased with increasing photoperiods for all spectral wavebands tested (Figure \@ref(fig:MIT9313PAR)), with the fastest overall growth rate for MIT9313 
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "WW", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under peak PAR of 90 µmol photons m^-2^ s^-1^ and 16 h white LED light photoperiod. In marked contrast to the 250 µM O~2~ growth experiments, MIT9313 grew when exposed to peak PAR of 180 µmol photons m^-2^ s^-1^ and blue light under all photoperiods except 16 h; additionally, white LED and red light treatments induced growth across all tested photoperiods under 25 µM O~2~. The 25 µM O~2~, 4 h photoperiod experiments under white LED light and were not performed due to time constraints.

MIT9313 grew under 2.5 µM O~2~  particularly under blue LED light, albeit generally slower than under the parallel experiments at 25 µM O~2~  (Figure \@ref(fig:MIT9313PAR)). Growth estimates showed scatter among replicates, suggesting  2.5 µM O~2~  is near the tolerance limit for growth of MIT9313. Growth rates increased with longer photoperiods under blue light treatments under peak PAR of 90 µmol photons m^-2^ s^-1^ . Growth  for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light treatment decreased with increasing photoperiods with full growth inhibition under a 16 h photoperiod. The red light peak PAR of 180 µmol photons m^-2^ s^-1^ showed similar growth rates for 8 and 12 h photoperiods. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 12, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 12 h blue light photoperiod and PAR of 90 µmol photons m^-2^ s^-1^. The 2.5 µM O~2~ white LED treatments under 4, 8 and 16 h photoperiods and red light under 4 and 16 h photoperiods were not performed due to time constraints.

@@

The GAM model in Figure \@ref(fig:MIT9313GAM) summarizes MIT9313 growth responses to red (A,B,C) or blue (D,E,F) peak PAR and photoperiod. Under 250 µM O~2~, Figure \@ref(fig:MIT9313GAM)D shows MIT9313 achieves fastest growth rates between blue peak PAR of 30 µmol photons m^-2^ s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 8 h, indicated by the contour line labeled 0.52 d^-1^ representing the 90^th^ percentile of achieved growth rates. Figure \@ref(fig:MIT9313GAM)D also shows that growth rate increases with longer photoperiods, as long as the blue peak PAR levels remain below 50 µmol photons m^-2^ s^-1^.  In contrast, under red light and 250 µM O~2~ MIT9313 grows faster while exploiting higher peak PAR and longer photoperiods.  Figure \@ref(fig:MIT9313GAM)F shows that MIT9313 maintains growth even under 2.5 µM O~2~, under photoperiods between 4 and 8 h and peak blue PAR between 50 to 10 µmol photons m^-2^ s^-1^ PAR.  Thus the designation of MIT9313 as a LL clade is dependent upon [O~2~] and light spectra. (Figure \@ref(fig:MIT9313GAM)E).

@@

Cumulative diel PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of photosynthetically active light absorbed per day. Cumulative diel PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed PUR (µmol photons m^-2^ s^-1^) and photoperiod (h). We plotted growth rates vs. cumulative diel PUR to determine whether growth is a simple response to diel PUR, across imposed photoperiods and spectral wavebands, or whether spectral wavebands or photoperiods have specific or interactive influences on growth beyond cumulative diel PUR.

Due to the absorption of *P. marinus* pigments in the blue spectral waveband range, the maximum imposed cumulative diel PUR under blue light is almost 3 times that of white LED light, and about 6 times the red light treatment, despite being derived from the same imposed photoperiods and peak PAR regimes. As such, only blue light experiments extend beyond cumulative diel PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^. This bias in the range of data leads to caution in comparing model fits of growth in response cumulative diel PUR under red, vs. blue, wavebands.

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth under blue light was well described by a saturating response of growth [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] to increasing cumulative diel blue PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:BluevsRedPurFitsPlots)A), and no evidence of inhibition of growth at any achieved cumulative diel PUR.  Under the 'artificial' growth treatment of red light, MED4 achieved more growth per unit diel cumulative PUR  \@ref(fig:BluevsRedPurFitsPlots)A), consistent with [@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for growth under red light, for MED4, because red light provokes less photoinactivation of PSII, than equivalent levels of blue light. For distinct photoperiod fits refer to figure \@ref(fig:AllStrainPurFitsPlots)A.

>>>>

Under 250 µM O~2~, growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and across all spectral wavebands (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate, however increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^ but showed growth inhibition at 16 h red light photoperiod.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white LED, red or blue light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^ growth trends were similar to 250 µM O~2~. SS120 showed no growth under a 4 h photoperiod for red spectral waveband, however under blue light, SS120 was able to grow (Figure \@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew under 4 and 8 h blue light and 16 h red light photoperiods, however the growth rate decreased under 12 and 16 h white LED light photoperiod treatments. Blue light treatments under PAR of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h photoperiod. The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

@@

The GAM model in Figure \@ref(fig:SS120GAM) summarizes growth responses of SS120 to red (A,B) or blue (C,D) peak PAR and photoperiod, across the 2 imposed oxygen concentrations. Under 250 µM O~2~, Figure \@ref(fig:SS120GAM)C showed highest growth rates below blue light PAR of 50 µmol photons m^-2^ s^-1^ and photoperiods between 8 and 12 h, indicated by the contour line labeled 0.19 d^-1^ (representing the 90^th^ percentile of achieved growth rate).  Under 250 µM O~2~ SS120 is constrained to deeper ocean waters through its intolerance of higher blue PAR levels. These findings align with Moore *et al*. [@mooreComparativePhysiologySynechococcus1995] and are expected for a low light clade.  The disjunct regions of the GAM plot results from variable growth success of SS120 under 250 µM O~2~.  Growth rate patterns under red light were similar, although somewhat faster. In contrast, under 25 µM O~2~ and a photoperiod of 8 h SS120 exploited all blue peak PAR levels, achieving faster growth rates at a higher PAR of \~100 µmol photons m^-2^ s^-1^, indicated by the contour line labeled 0.4 d^-1^ (representing the 90^th^ percentile of achieved growth rate), out pacing the 90^th^ percentile fastest growth rates under 250 µM O~2~ (Figure \@ref(fig:SS120GAM)D). Under red light and 25 µM O~2~ (Figure \@ref(fig:SS120GAM)B) SS120 grew across most conditions of peak PAR and photoperiod, achieving fastest growth under long photoperiods and peak PAR between 50 \~100 µmol photons m^-2^ s^-1^.  Thus, the designation of SS120 as a LL strain is dependent upon the [O~2~]. SS120 did not, however, grow reliably under tested conditions at 2.5 µM O~2~.

@@

Under 250 µM O~2~ growth rates of *P. marinus* MIT9313, Clade LLIV, increased with longer
photoperiods, under low 30 µmol photons m^-2^ s^-1^ peak PAR, (Figure \@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiods. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light at 180 µmol photons m^-2^ s^-1^ peak PAR, under the 8 h photoperiod, consistent with Moore *et al*. [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

For MIT9313 under 25 µM O~2~, growth rate increased with increasing photoperiods for all spectral wavebands tested (Figure \@ref(fig:MIT9313PAR)), with the fastest overall growth rate for MIT9313 
`r PURPARFits %>% filter(Strain == "MIT9313", WL == "WW", O2_uM == "25", Photoperiod == 16, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under peak PAR of 90 µmol photons m^-2^ s^-1^ and 16 h white LED light photoperiod. In marked contrast to the 250 µM O~2~ growth experiments, MIT9313 grew when exposed to peak PAR of 180 µmol photons m^-2^ s^-1^ and blue light under all photoperiods except 16 h; additionally, white LED and red light treatments induced growth across all tested photoperiods under 25 µM O~2~. The 25 µM O~2~, 4 h photoperiod experiments under white LED light and were not performed due to time constraints.

MIT9313 grew under 2.5 µM O~2~  particularly under blue LED light, albeit generally slower than under the parallel experiments at 25 µM O~2~  (Figure \@ref(fig:MIT9313PAR)). Growth estimates showed scatter among replicates, suggesting  2.5 µM O~2~  is near the tolerance limit for growth of MIT9313. Growth rates increased with longer photoperiods under blue light treatments and peak PAR of 90 µmol photons m^-2^ s^-1^ but did not grow under 16 h photoperiod. Growth  for MIT9313 under PAR of 180 µmol photons m^-2^ s^-1^ and blue light treatment decreased with increasing photoperiods with full growth inhibition under a 16 h photoperiod. The red light peak PAR of 180 µmol photons m^-2^ s^-1^ showed similar growth rates to blue light for 8 and 12 h photoperiods. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "2.5", Photoperiod == 12, Par_ue == 90) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 12 h blue light photoperiod and PAR of 90 µmol photons m^-2^ s^-1^. The 2.5 µM O~2~ white LED treatments under 4, 8 and 16 h photoperiods and red light under 4 and 16 h photoperiods were not performed due to time constraints.

@@

The GAM model in Figure \@ref(fig:MIT9313GAM) summarizes MIT9313 growth responses to red (A,B,C) or blue (D,E,F) peak PAR and photoperiod. Under 250 µM O~2~, Figure \@ref(fig:MIT9313GAM)D shows MIT9313 achieves fastest growth rates between blue peak PAR of 30 µmol photons m^-2^ s^-1^ and 50 µmol photons m^-2^ s^-1^ and photoperiods longer than 8 h, indicated by the contour line labeled 0.52 d^-1^ representing the 90^th^ percentile of achieved growth rates. Figure \@ref(fig:MIT9313GAM)D also shows that growth rate increases with longer photoperiods, as long as the blue peak PAR levels remain below 50 µmol photons m^-2^ s^-1^.  In contrast, under red light and 250 µM O~2~ MIT9313 grows faster while exploiting higher peak PAR and longer photoperiods. Figure \@ref(fig:MIT9313GAM)E shows that MIT9313can exploit all blue PAR levels and most photoperiods with 90^th^ percentile of fastest growth rate between 30 to 100 µmol photons m^-2^ s^-1^ PAR.F igure \@ref(fig:MIT9313GAM)F shows that MIT9313 maintains growth even under 2.5 µM O~2~, under photoperiods between 4 and 8 h and peak blue PAR between 50 to 100 µmol photons m^-2^ s^-1^ PAR.  Thus the designation of MIT9313 as a LL clade is dependent upon [O~2~] and light spectra. (Figure \@ref(fig:MIT9313GAM)E).

@@

Cumulative diel PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of photosynthetically active light absorbed per day. Cumulative diel PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed PUR XXX should say imposed PAR?? or absorbed PUR ?? XXX (µmol photons m^-2^ s^-1^) and photoperiod (h). We plotted growth rates vs. cumulative diel PUR to determine whether growth is a simple response to diel PUR, across imposed photoperiods and spectral wavebands, or whether spectral wavebands or photoperiods have specific or interactive influences on growth beyond cumulative diel PUR.

Due to the absorption of *P. marinus* pigments in the blue spectral waveband range, the maximum imposed cumulative diel PUR under blue light is almost 3 times that of white LED light, and about 6 times the red light treatment (Figure \@ref(fig:PurParPlot)), despite being derived from the same imposed photoperiods and peak PAR regimes. As such, only blue light experiments extend beyond cumulative diel PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^. This bias in the range of data leads to caution in comparing model fits of growth in response to cumulative diel PUR under red vs. blue wavebands.

@@

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR equivalent to other photoperiod treatments. In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth under blue light was well described by a saturating response of growth [@harrisonPhotosynthesisirradianceRelationshipsPolar1986] to increasing cumulative diel PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:BluevsRedPurFitsPlots)A), and no evidence of inhibition of growth at any achieved cumulative diel PUR.  Under the 'artificial' growth treatment of red light, MED4 achieved more growth per unit diel cumulative PUR (Figure \@ref(fig:BluevsRedPurFitsPlots)A), consistent with Murphy *et al*. [@murphyPhotoinactivationPhotosystemII2017], who showed a lower cost for growth under red light, for MED4, because red light provokes less photoinactivation of PSII, than equivalent levels of blue light. For distinct photoperiod fits refer to figure \@ref(fig:AllStrainPurFitsPlots)A.
